Challenges to the Two-Infall Scenario by Large Stellar Age Catalogs

### 2025-05-16 src/tex/ms.tex

% user-defined commands
\newcommand{\vice}{{\tt VICE}\xspace}
\newcommand{\todo}[1]{{\color{red}#1}}
\newcommand{\osuaffil}{Department of Astronomy, The Ohio State University, 140 W. 18th Ave, Columbus OH 43210, USA}
\newcommand{\ccappaffil}{Center for Cosmology and AstroParticle Physics, The Ohio State University, 191 W. Woodruff Ave., Columbus OH 43210, USA}
\newcommand{\aFe}{[$\alpha$/Fe]\xspace}
\newcommand{\mathXH}{{\rm [X/H]}}
\newcommand{\mathOH}{{\rm [O/H]}}
\newcommand{\mathFeH}{{\rm [Fe/H]}}
\newcommand{\mathOFe}{{\rm [O/Fe]}}
\newcommand{\mathaFe}{[\alpha/{\rm Fe}]}
\newcommand{\yZ}[1]{$y/Z_\odot=#1$}
\newcommand{\kpc}{\,{\rm kpc}}
\newcommand{\Gyr}{\,{\rm Gyr}}

@@

% Title
\title{Limitations on the Two-Infall Scenario in the Era of Large Stellar Age Catalogs}

@@

\begin{abstract}
    Stellar populations in the Milky Way exhibit a clear separation into two chemically distinct populations that are separated by their [$\alpha$/Fe] ratios. This $\alpha$-bimodality is not a universal feature of simulated disk galaxies and may point to the Milky Way's unique evolutionary history. A popular explanation for this observation is the two-infall scenario, which postulates that two periods of substantial accretion rates dominate the assembly history of the Galaxy. However, most previous studies using the two-infall scenario have explored a limited portion of the parameter space, typically neglecting radial migration and assuming that the Galactic disk never ejected a substantial outflow. Thanks to advances in stellar age measurements in recent years, we can now also compare this popular model to more direct measurements of the Galaxy's evolutionary timescales across the disk from large stellar catalogs. We run multi-zone galactic chemical evolution (GCE) models with a two-infall-driven star formation history, radially dependent mass-loaded outflows, and a prescription for radial migration tuned to a hydrodynamic simulation. We compare our model results to abundance patterns across the disk from APOGEE DR17,
    % and early data from the SDSS-V Milky Way Mapper program,
    supplemented with stellar age estimates through multiple methods. Although the two-infall scenario offers a natural explanation for the [$\alpha$/Fe] bimodality, it struggles to explain several features of the age--abundances structure in the disk. The two-infall scenario generically predicts a massive and long-lasting dilution event, but the data show that stellar metallicity is remarkably constant with age across much of the Galactic disk. This apparent age-independence places considerable restrictions upon the two-infall parameter space. These issues can be mitigated, but not completely resolved, by allowing the accreted gas to be pre-enriched to low metallicity.
\end{abstract}

@@

% \todo{Hook, summary of alpha-bimodality.}
\lettrine{G}{alactic chemical evolution} (GCE) studies aim to explain the observed stellar abundance patterns in the Milky Way (MW) by modeling the star formation history and evolution of the Galaxy. A long-standing paradigm of GCE is that the metallicity of the interstellar medium (ISM) increases over time due to supernova enrichment from successive generations of stars. In this view, one feature of the MW disk that is difficult to explain is the so-called ``$\alpha$-bimodality'': the presence of two populations of stars at similar metallicity but separated by their \aFe ratio \citep[e.g.,][]{bensby_exploring_2014}. The high-$\alpha$ sequence consists of old stars \citep[$\gtrsim9$ Gyr;][]{pinsonneault_apokasc-3_2025} with super-Solar \aFe and is associated with the kinematic thick disk \citep[e.g.,][]{fuhrmann_nearby_1998}, while the low-$\alpha$ sequence is younger, with approximately Solar \aFe, and is associated with the thin disk. The $\alpha$-bimodality is present across the Galactic disk, but the relative strength of the high- and low-$\alpha$ sequences varies by location \citep{hayden_chemical_2015}.

@@

% Summary of two-infall literature
The two-infall model of chemical evolution was proposed by \citet{chiappini_chemical_1997} to explain the origin of the high- and low-$\alpha$ disks. Though the model has been revised and refined since, the basic premise remains the same: the rate of gas infall onto the Galaxy is described by two consecutive, exponentially declining bursts. The relatively low infall rate between the two bursts leads to a hiatus in the star formation rate, allowing the gas abundance to evolve between the high- and low-$\alpha$ sequences while producing few stars in between. 
The infall timescale for the low-$\alpha$ disk can be varied to produce inside-out disk growth and a radial metallicity gradient \citep{romano_mass_2000}.
The initial model of \citet{chiappini_chemical_1997} successfully reproduced the available abundance data at the time, which were largely confined to the Solar neighborhood.

@@

% \todo{Migration.} 
The two-infall model attempts to reproduce the full distribution of stellar abundances in the Solar neighborhood through a single, continuous evolutionary track. However, the current body of evidence suggests that many of the stars that make up the wings of the local metallicity distribution originate from elsewhere in the Galaxy. \citet{sellwood_radial_2002} first showed that transient spiral perturbations can induce large changes in the guiding radius of a star without kinematic heating, and it is now understood that the stars that make up the Solar neighborhood are drawn from a wide range of birth radii \citep[e.g.,][]{schonrich_chemical_2009,frankel_measuring_2018,lehmann_probing_2024}. Some studies have attempted to derive stellar birth radii \citep[e.g.,][]{ratcliffe_unveiling_2023,lu_there_2024}, though such an endeavor requires assumptions about the evolution of the Milky Way's radial metallicity gradient. While the strength and speed of radial migration in the disk are not precisely measured, it is clear that a single chemical evolution track need not explain the entirety of the local observed abundance distribution.

% \todo{Observational evidence other than $\alpha$-bimodality.} 
The chemical and kinematic separation of the high- and low-$\alpha$ disks remains the primary observational evidence behind the two-infall model. \citet{spitoni_remind_2024} argued that the observed gap between the sequences in \aFe, contrasted with their overlap in $[\alpha/{\rm H}]$, indicates a haitus in the star formation history, which is a natural consequence of the two-infall model. In different samples, \citet{nissen_high-precision_2020} and \citet{nataf_accurate_2024} observed multiple sequences in the local age--metallicity relation, which would naturally be explained by the two-infall scenario. Many two-infall studies have also reproduced the metallicity gradient, the local surface densities of stars and gas, and the local star formation and supernova rates \todo{(citations)}, although these observables can also be matched by other models \todo{(citations)}.

@@

We supplement the APOGEE DR17 abundance data with two different age catalogs. The first is from \citet{leung_variational_2023}, who train a variational encoder-decoder network on asteroseismic data for APOGEE red giants. This catalog has two main advantages over other neural network (NN) age estimates: their method is designed to reduce contamination from abundance information (in particular [$\alpha$/Fe]), and the recovered ages do not plateau at $\sim10\,{\rm Gyr}$ as they do in some other neural network-derived age catalogs \citep[e.g.,][]{mackereth_dynamical_2019}. Following the recommendations of \citet{leung_variational_2023}, we use an upper limit on the relative age uncertainty of 40\%. This produces a sample of \num{57607} stars with NN age estimates, of which \num{14871} are in the Solar neighborhood ($7\leq R_{\rm gal}<9\kpc$, $0\leq|z|<0.5\kpc$). The median uncertainty in $\log({\rm age}/{\rm Gyr})$ is 0.10 (see Table \ref{tab:uncertainties}). We hereafter refer to the \citet{leung_variational_2023} ages as ``NN ages.''

@@

We discuss our assumptions about the nucleosynthetic yields in Section \ref{sec:yields}, and we detail our outflow prescription in Section \ref{sec:outflows}. We describe the gas supply in Section \ref{sec:sfh}, we detail the infall parameter selection in Section \ref{sec:parameter-selection}, and we present our star formation law in Section \ref{sec:sf-law}. We briefly summarize the stellar migration prescription in Section \ref{sec:migration}. Table \ref{tab:multizone-parameters} summarizes the most relevant variables and their fiducial values in this work.

@@

The population-averaged nucleosynthetic yields of CCSNe are still fairly uncertain. This problem is exacerbated by the complexity of the CCSN explosion landscape \citep{sukhbold_core-collapse_2016}. Recently, \citet{weinberg_scale_2024} used a measurement of the mean Fe yield of CC SNe by \citet{rodriguez_iron_2023} and the plateau in stellar \aFe abundances at low metallicity to infer population-averaged yields of $y/Z_\odot\approx1$\textemdash in other words, the scale of SN yields is similar to the Solar abundance. However, \citet{johnson_milky_2024} found that GCE models with yields at this scale approach present-day abundances too slowly to match the observed age--metallicity relation. Previous multi-zone models using {\tt VICE} \citep[e.g.,][]{johnson_stellar_2021,dubay_galactic_2024} adopted higher yields ($y/Z_\odot\approx2.6$) based on \citet{chieffi_explosive_2004} and \citet{limongi_nucleosynthesis_2006}; however, in order to produce a realistic evolution of [O/Fe], those studies adopted an integrated SN Ia rate which is high compared to the measurement of \citet{maoz_star_2017}.

We investigate yield sets at multiple scales of the Solar abundance. The CCSN yield of O is directly set by the Solar scale: $y_{\rm O}^{\rm CC}=(y/Z_\odot)Z_{\rm O,\odot}$. For Fe, the CCSN yield is set by the \aFe ``plateau'' at low metallicity, $\mathaFe_{\rm CC}$, such that $y_{\rm Fe}^{\rm CC}=(y/Z_\odot) Z_{\rm Fe,\odot} 10^{-\mathaFe}$ \citep[for further discussion on the empirical yield scale and the CCSN plateau, see][]{weinberg_scale_2024}. We adopt the \citet{asplund_chemical_2009} Solar abundances: $Z_{\rm O,\odot}=5.72\times10^{-3}$ and $Z_{\rm Fe,\odot}=1.29\times10^{-3}$. Our yield sets are presented in Table \ref{tab:yields}. We consider \yZ{1} representative of the empirical yield scale, whereas \yZ{2-3} span a range of theoretical predictions.

The SN Ia yield of Fe, $y_{\rm Fe}^{\rm Ia}$, is calculated so that the combined Fe yield $y_{\rm Fe}^{\rm Ia}+y_{\rm Fe}^{\rm CC}=y/Z_\odot$. We further increase $y_{\rm Fe}^{\rm Ia}$ by a factor of 30\% and 10\% for the \yZ{1} and \yZ{2} yield sets, respectively, to ensure that our models reach $\mathOFe\approx0.0$ by $t=13.2\,{\rm Gyr}$. The fifth row of Table \ref{tab:yields} reports the integrated SN Ia rate 
\begin{equation}
    \frac{N_{\rm Ia}}{M_\star} = \frac{y_{\rm Fe}^{\rm Ia}}{\bar m_{\rm Fe}^{\rm Ia}}
    \label{eq:snia-rate}
\end{equation}
from each yield set, assuming a mean Fe yield per SN Ia of $\overline m_{\rm Fe}^{\rm Ia}=0.7$ M$_\odot$ \citep{mazzali_common_2007,howell_effect_2009}. The rate for the \yZ{1} yield set is slightly higher than the volumetric rate of $N_{\rm Ia}/M_\star=(1.3\pm0.1)\times10^{-3}\,{\rm M}_\odot^{-1}$ reported by \citet{maoz_star_2017}, but is consistent with their measurement of $N_{\rm Ia}/M_\star=(1.6\pm0.3)\times10^{-3}\,{\rm M}_\odot^{-1}$ for field galaxies. The rate for the \yZ{2} yield set is consistent with the measurement of $N_{\rm Ia}/M_\star=(2.2\pm1.0)\times10^{-3}\,{\rm M}_\odot^{-1}$ by \citet{maoz_type-ia_2012}, while the rate for the \yZ{3} yield set is generally higher than literature values.

@@

The three models in Figure \ref{fig:yield-outflow} predict nearly identical evolution in [O/Fe] over the past 5 Gyr, and the trend in the data is similar apart from a $\sim0.05$ dex offset (which could be corrected by raising the SN Ia yield). The offset between the data and models grows between $\sim5-8\,{\rm Gyr}$, especially for the higher yields. Of the three, the \yZ{2} and \yZ{3} models most closely match the observed late-time age--metallicity relations, whereas the \yZ{1} model shows the best agreement with the observed [O/Fe] evolution. As the \yZ{2} and \yZ{3} models behave qualitatively similarly, we focus on the \yZ{1} and \yZ{2} yield sets for the remainder of this study.

@@

Mass-loaded outflows are a useful tool for scaling the endpoint of GCE models. \citet{weinberg_equilibrium_2017} showed that in the case of exponentially declining star formation, the O abundance approaches an equilibrium value at
\begin{equation}
    \label{eq:equilibrium}
    Z_{\rm O,eq} = \frac{y_{\rm O}^{\rm CC}}{1 + \eta - r - \tau_\star/\tau_{\rm SFH}},
\end{equation}
where $r=0.4$ is the instantaneous recycling parameter, $\tau_\star$ is the star formation efficiency timescale, $\tau_{\rm SFH}$ is the star formation timescale, and $\eta\equiv \dot\Sigma_{\rm out}/\dot\Sigma_\star$ is the outflow mass-loading factor. Equation \ref{eq:equilibrium} can be inverted to calculate the value of $\eta$ required to reach a desired $Z_{\rm O,eq}$ with a given yield, but it is inaccurate in the limit of zero accretion. We adopt values of $\eta_\odot=\eta(R=R_\odot)$ for each of our yield sets presented in Table \ref{tab:yields}. Models with lower yields do not achieve a steady-state abundance in time (see Figure \ref{fig:yield-outflow}); therefore, the values of $\eta_\odot$ for \yZ{1} and \yZ{2} are lower than would be suggested by Equation \ref{eq:equilibrium} in order to reach Solar metallicity at the end of the model.

@@

We run {\tt VICE} in ``infall mode,'' where we specify the gas infall density $\dot\Sigma_{\rm in}$ and the star formation effieciency (SFE) timescale $\tau_\star\equiv \Sigma_g / \dot\Sigma_\star$ as functions of time. The gas surface density $\Sigma_g$ and star formation rate $\dot\Sigma_\star$ are calculated from the two specified quantities according to our star formation law, which is described in Section \ref{sec:sf-law}, assuming zero initial gas mass in all zones.

@@

Panel (b) of Figure \ref{fig:twoinfall-parameters} shows that 
% the second infall timescale $\tau_2$ is the most important parameter to tune in order to reproduce the observed stellar abundance distributions. 
the second infall timescale $\tau_2$ controls the width of the MDF and the low-$\alpha$ [O/Fe] distribution. A shorter $\tau_2$ produces a broader [O/Fe] distribution which is skewed to higher [O/Fe], while a longer $\tau_2$ produces both a narrower low-$\alpha$ sequence and a narrower MDF. We note that our maximum value of $\tau_2=30\,{\rm Gyr}$ is very close to a constant infall rate, so a further increase in $\tau_2$ has diminishing returns. Between $\tau_2=3-30\,{\rm Gyr}$, the endpoint of the abundance tracks shifts by $\sim0.2$ dex in [Fe/H] and $\sim0.1$ dex in [O/Fe], which could affect the model's ability to reproduce the present-day abundance of the Solar neighborhood. We adopt a fiducial value of $\tau_2=15\,{\rm Gyr}$ for the Solar neighborhood in order to minimize the width of the low-$\alpha$ [O/Fe] distribution while still approaching Solar [Fe/H] at late times. This value is in line with the infall timescale recovered by \citet{spitoni_galactic_2020}, and similar to the local star formation timescale adopted by \citet{johnson_stellar_2021}, but significantly longer than the timescales found by \citet{nissen_high-precision_2020} and \citet{spitoni_apogee_2021}. 
% In \todo{Section X}, we explore the effect of varying $\tau_2$ with radius in multi-zone models.

@@

We note that all our models are normalized to produce the same thick-to-thin-disk mass ratio of $f_{\Sigma,0}=0.27$ \citep{bland-hawthorn_galaxy_2016} at the Solar annulus regardless of the infall parameters. The high-$\alpha$ sequence appears much less prominent in our [O/Fe] distributions in Figure \ref{fig:twoinfall-parameters} than in the data because the model outputs include only stars which were formed in-situ at the Solar annulus. In our multi-zone models, most of the high-$\alpha$ stars present in the Solar neighborhood have migrated from the inner Galaxy.

@@

\begin{figure*}
    \centering
    \includegraphics{figures/abundance_evolution_params.pdf}
    \caption{Stellar age--abundance relations produced by select multi-zone models with \yZ{1}. The layout is similar to Figure \ref{fig:abundance-evolution-yields}. Each column shows results from a different multi-zone model: {\bf (a)} our fiducial model with $y/Z_\odot=1$, $\sigma_{\rm RM8}=2.7\kpc$, pristine gas infall, and $f_\Sigma(R_\odot)=0.12$; {\bf (b)} a model with greater radial migration strength $\sigma_{\rm RM8}=5\kpc$; {\bf (c)} a model that assumes the infalling gas has metallicity $\mathOH_{\rm CGM}=\mathFeH_{\rm CGM}=-0.5$; and {\bf (d)} a model with a higher local thick-to-thin disk ratio, $f_\Sigma(R_\odot)=0.5$. {\bf Key takeaway:} Model (d) comes the closest to the observed age--metallicity relation at the expense of the age--[O/Fe] relation, but no model completely reconciles the dilution problem.}
    \label{fig:abundance-evolution-params}
\end{figure*}

>>>>

% user-defined commands
\newcommand{\vice}{{\tt VICE}\xspace}
\newcommand{\todo}[1]{{\color{red}#1}}
\newcommand{\osuaffil}{Department of Astronomy, The Ohio State University, 140 W. 18th Ave, Columbus OH 43210, USA}
\newcommand{\ccappaffil}{Center for Cosmology and AstroParticle Physics, The Ohio State University, 191 W. Woodruff Ave., Columbus OH 43210, USA}
\newcommand{\aFe}{[$\alpha$/Fe]\xspace}
\newcommand{\mathXH}{{\rm [X/H]}}
\newcommand{\mathOH}{{\rm [O/H]}}
\newcommand{\mathFeH}{{\rm [Fe/H]}}
\newcommand{\mathOFe}{{\rm [O/Fe]}}
\newcommand{\mathaFe}{[\alpha/{\rm Fe}]}
\newcommand{\yZ}[1]{$y/Z_\odot=#1$}
\newcommand{\kpc}{\,{\rm kpc}}
\newcommand{\Gyr}{\,{\rm Gyr}}
\newcommand{\Msun}{\,{\rm M}_\odot}

@@

% Title
\title{Challenges to the Two-Infall Scenario by Large Stellar Age Catalogs}

@@

\begin{abstract}
    Stellar populations in the Milky Way exhibit a clear separation into two chemically distinct populations that are separated by their [$\alpha$/Fe] ratios. This $\alpha$-bimodality is not a universal feature of simulated disk galaxies and may point to the Milky Way's unique evolutionary history. A popular explanation for this observation is the two-infall scenario, which postulates that two periods of substantial accretion rates dominate the assembly history of the Galaxy. However, most previous studies using the two-infall scenario have explored a limited portion of the parameter space, typically neglecting radial migration and assuming that the Galactic disk never ejected a substantial outflow. Thanks to advances in stellar age measurements in recent years, we can now also compare this popular model to more direct measurements of the Galaxy's evolutionary timescales across the disk from large stellar catalogs. We run multi-zone galactic chemical evolution (GCE) models with a two-infall-driven star formation history, radially dependent mass-loaded outflows, and a prescription for radial migration tuned to a hydrodynamic simulation. We compare our model results to abundance patterns across the disk from APOGEE DR17,
    % and early data from the SDSS-V Milky Way Mapper program,
    supplemented with stellar age estimates through multiple methods. Although the two-infall scenario offers a natural explanation for the [$\alpha$/Fe] bimodality, it struggles to explain several features of the age--abundances structure in the disk. The two-infall scenario generically predicts a massive and long-lasting dilution event, but the data show that stellar metallicity is remarkably constant with age across much of the Galactic disk. This apparent age-independence places considerable restrictions upon the two-infall parameter space. These issues can be mitigated, but not completely resolved, by allowing the accreted gas to be pre-enriched to low metallicity. \todo{Add a sentence about larger implications for merger-dominated SFHs.}
\end{abstract}

@@

% \todo{Hook, summary of alpha-bimodality.}
\lettrine{G}{alactic chemical evolution} (GCE) studies aim to explain the observed stellar abundance patterns in the Milky Way (MW) by modeling the star formation history and evolution of the Galaxy. A long-standing paradigm of GCE is that the metallicity of the interstellar medium (ISM) increases over time due to supernova enrichment from successive generations of stars. In this view, one feature of the MW disk that is difficult to explain is the so-called ``$\alpha$-bimodality'': the presence of two populations of stars at similar metallicity but separated by their \aFe ratio \citep[e.g.,][]{bensby_exploring_2014}. The high-$\alpha$ sequence consists of old stars \citep[$\gtrsim9$ Gyr; e.g.,][]{pinsonneault_apokasc-3_2025} with super-Solar \aFe and is associated with the kinematic thick disk \citep[e.g.,][]{fuhrmann_nearby_1998}, while the low-$\alpha$ sequence is younger, with approximately Solar \aFe, and is associated with the thin disk. The $\alpha$-bimodality is present across the Galactic disk, but the relative strength of the high- and low-$\alpha$ sequences varies by location \citep{hayden_chemical_2015}.

@@

% Summary of two-infall literature
The two-infall model of chemical evolution was proposed by \citet{chiappini_chemical_1997} to explain the origin of the high- and low-$\alpha$ disks. Though the model has been revised and refined since, the basic premise remains the same: the rate of gas infall onto the Galaxy is described by two consecutive, exponentially declining bursts. The relatively low infall rate between the two bursts leads to a lower star formation rate, allowing the gas abundance to evolve between the high- and low-$\alpha$ sequences while producing few stars in between. 
The infall timescale for the low-$\alpha$ disk can be varied to produce inside-out disk growth and a radial metallicity gradient \citep{romano_mass_2000}.
The initial model of \citet{chiappini_chemical_1997} successfully reproduced the available abundance data at the time, which were largely confined to the Solar neighborhood.

@@

\todo{Discuss yields, degeneracy with outflows and Rodriguez measurements. Emphasize that degeneracy with outflows prohibits direct estimates of yields from GCE. We aim to study the effect of the yield assumptions, which necessitates adjusting the outflows as well.}

% \todo{Migration.} 
The two-infall model attempts to reproduce the full distribution of stellar abundances in the Solar neighborhood through a single, continuous evolutionary track. However, the current body of evidence suggests that many of the stars that make up the wings of the local metallicity distribution originate from elsewhere in the Galaxy. \citet{sellwood_radial_2002} first showed that transient spiral perturbations can induce large changes in the guiding radius of a star without kinematic heating, and it is now understood that the stars that make up the Solar neighborhood are drawn from a wide range of birth radii \citep[e.g.,][]{schonrich_chemical_2009,frankel_measuring_2018,lehmann_probing_2024}. Some studies have attempted to derive stellar birth radii \citep[e.g.,][]{ratcliffe_unveiling_2023,lu_there_2024}, though such an endeavor requires also reconstructing the evolution of the Milky Way's radial metallicity gradient. While the strength and speed of radial migration in the disk are not precisely measured, it is clear that a single chemical evolution track need not explain the entirety of the local observed abundance distribution.

% \todo{Observational evidence other than $\alpha$-bimodality.} 
The chemical and kinematic separation of the high- and low-$\alpha$ disks remains the primary observational evidence behind the two-infall model. \citet{spitoni_remind_2024} argued that the observed gap between the sequences in \aFe, contrasted with their overlap in $[\alpha/{\rm H}]$, indicates a period of reduced star formation, which is a natural consequence of the two-infall model. In different samples, \citet{nissen_high-precision_2020} and \citet{nataf_accurate_2024} observed multiple sequences in the local age--metallicity relation, which would naturally be explained by the two-infall scenario. Many two-infall studies have also reproduced the metallicity gradient, the local surface densities of stars and gas, and the local star formation and supernova rates \todo{(citations)}, although these observables can also be matched by other models \todo{(citations)}.

@@

We supplement the APOGEE DR17 abundance data with two different age catalogs. The first is from \citet{leung_variational_2023}, who train a variational encoder-decoder network on asteroseismic data for APOGEE red giants. This catalog has two main advantages over other neural network (NN) age estimates: their method is designed to reduce contamination from abundance information (in particular [$\alpha$/Fe]), and the recovered ages do not plateau at $\sim10\,{\rm Gyr}$ as they do in some other neural network-derived age catalogs \citep[e.g.,][]{mackereth_dynamical_2019}. Following the recommendations of \citet{leung_variational_2023}, we cut all stars which have a relative age uncertainty greater than 40\%. This produces a sample of \num{57607} stars with NN age estimates, of which \num{14871} are in the Solar neighborhood ($7\leq R_{\rm gal}<9\kpc$, $0\leq|z|<0.5\kpc$). The median uncertainty in $\log({\rm age}/{\rm Gyr})$ is 0.10 (see Table \ref{tab:uncertainties}). We hereafter refer to the \citet{leung_variational_2023} ages as ``NN ages.''

@@

We discuss our assumptions about the nucleosynthetic yields in Section \ref{sec:yields}, the outflow prescription in Section \ref{sec:outflows}, the gas supply in Section \ref{sec:sfh}, the infall parameter selection in Section \ref{sec:parameter-selection}, the star formation law in Section \ref{sec:sf-law}, and the stellar migration prescription in Section \ref{sec:migration}. Table \ref{tab:multizone-parameters} summarizes the most relevant variables and their fiducial values in this work.

@@

The population-averaged nucleosynthetic yields of CCSNe, $y_{\rm X}^{\rm CC}$, are uncertain to a degree that is significant for chemical evolution models. This problem is exacerbated by the complexity of the CCSN explosion landscape \citep{sukhbold_core-collapse_2016}. Recently, \citet{weinberg_scale_2024} used a measurement of the mean Fe yield of CC SNe by \citet{rodriguez_iron_2023} and the plateau in stellar \aFe abundances at low metallicity to infer population-averaged yields of $y/Z_\odot\approx1$\textemdash in other words, for every $1\Msun$ of stars formed, massive stars release a mass of newly-synthesized $\alpha$-elements (e.g., O or Mg) equal to their mass in the Sun. However, \citet{johnson_milky_2024} found that GCE models with yields at this scale approach present-day abundances too slowly to match the observed age--metallicity relation. Previous multi-zone models using {\tt VICE} \citep[e.g.,][]{johnson_stellar_2021,dubay_galactic_2024} adopted higher yields ($y/Z_\odot\approx2.6$) based on \citet{chieffi_explosive_2004} and \citet{limongi_nucleosynthesis_2006}; however, in order to produce a realistic evolution of [O/Fe], those studies adopted an integrated SN Ia rate which is high compared to the measurement of \citet{maoz_star_2017}.

We investigate yield sets at multiple scales of the Solar abundance. The CCSN yield of O is directly set by the Solar scale, $y_{\rm O}^{\rm CC}=(y/Z_\odot)Z_{\rm O,\odot}$, because all O is assumed to form in CCSNe. For Fe, the CCSN yield is set by the \aFe ``plateau'' at low metallicity, $\mathaFe_{\rm CC}$, such that $y_{\rm Fe}^{\rm CC}=(y/Z_\odot) Z_{\rm Fe,\odot} 10^{-\mathaFe}$ \citep[for further discussion on the empirical yield scale and the CCSN plateau, see][]{weinberg_scale_2024}. We adopt the \citet{asplund_chemical_2009} Solar abundances: $Z_{\rm O,\odot}=5.72\times10^{-3}$ and $Z_{\rm Fe,\odot}=1.29\times10^{-3}$. Our yield sets are presented in Table \ref{tab:yields}. We consider \yZ{1} representative of the empirical yield scale, whereas \yZ{2-3} span a range of theoretical predictions.

The SN Ia yield of Fe, $y_{\rm Fe}^{\rm Ia}$, is set so that our models reach $\mathOFe\approx0.0$ by $t=13.2\,{\rm Gyr}$. For \yZ{3}, the combined Fe yield of CCSNe and SNe Ia matches the Solar yield scale: $(y_{\rm Fe}^{\rm Ia}+y_{\rm Fe}^{\rm CC})/Z_{\rm Fe,\odot}=y/Z_\odot$; for \yZ{1} and \yZ{2}, $y_{\rm Fe}^{\rm Ia}$ is increased from the given scale by a factor of 30\% and 10\%, respectively \todo{(rephrase)}. The fifth row of Table \ref{tab:yields} reports the integrated SN Ia rate 
\begin{equation}
    \frac{N_{\rm Ia}}{M_\star} = \frac{y_{\rm Fe}^{\rm Ia}}{\bar m_{\rm Fe}^{\rm Ia}}
    \label{eq:snia-rate}
\end{equation}
from each yield set, assuming a mean Fe yield per SN Ia of $\overline m_{\rm Fe}^{\rm Ia}=0.7$ M$_\odot$ \citep{mazzali_common_2007,howell_effect_2009}. The rate for the \yZ{1} yield set is slightly higher than the volumetric rate of $N_{\rm Ia}/M_\star=(1.3\pm0.1)\times10^{-3}\,{\rm M}_\odot^{-1}$ reported by \citet{maoz_star_2017}, but is consistent with their measurement of $N_{\rm Ia}/M_\star=(1.6\pm0.3)\times10^{-3}\,{\rm M}_\odot^{-1}$ for field galaxies. The rate for the \yZ{2} yield set is consistent with the measurement of $N_{\rm Ia}/M_\star=(2.2\pm1.0)\times10^{-3}\,{\rm M}_\odot^{-1}$ by \citet{maoz_type-ia_2012}, while the rate for the \yZ{3} yield set is generally higher than literature values.

@@

The three models in Figure \ref{fig:yield-outflow} predict nearly identical evolution in [O/Fe] over the past 5 Gyr, and the trend in the data is similar apart from a $\sim0.05$ dex offset (which could be corrected by increasing $y_{\rm Fe}^{\rm Ia}$ or adjusting the zero-point in the stellar abundances). The offset between the data and models grows between $\sim5-8\,{\rm Gyr}$, especially for the higher yields. Of the three, the \yZ{2} and \yZ{3} models most closely match the observed late-time age--metallicity relations, whereas the \yZ{1} model shows the best agreement with the observed [O/Fe] evolution. As the \yZ{2} and \yZ{3} models behave qualitatively similarly, we focus on the \yZ{1} and \yZ{2} yield sets for the remainder of this study.

@@

Mass-loaded outflows are a useful tool for scaling the endpoint of GCE models. \citet{weinberg_equilibrium_2017} showed that in the case of exponentially declining star formation, the O abundance approaches an equilibrium at
\begin{equation}
    \label{eq:equilibrium}
    Z_{\rm O,eq} = \frac{y_{\rm O}^{\rm CC}}{1 + \eta - r - \tau_\star/\tau_{\rm SFH}},
\end{equation}
where $r=0.4$ is the instantaneous recycling parameter, $\tau_\star$ is the star formation efficiency timescale, $\tau_{\rm SFH}$ is the star formation timescale, and $\eta\equiv \dot\Sigma_{\rm out}/\dot\Sigma_\star$ is the outflow mass-loading factor. 
% Equation \ref{eq:equilibrium} can be inverted to calculate the value of $\eta$ required to reach a desired $Z_{\rm O,eq}$ with a given yield, but it is inaccurate in the limit of zero accretion. 
Motivated by Equation \ref{eq:equilibrium}, we adopt an outflow mass-loading factor at the Solar radius $\eta_\odot\equiv\eta(R=R_\odot)$ for each of the yield sets in Table \ref{tab:yields}. Models with lower yields do not achieve a steady-state abundance in time (see Figure \ref{fig:yield-outflow}); therefore, the values of $\eta_\odot$ for \yZ{1} and \yZ{2} are lower than would be suggested by Equation \ref{eq:equilibrium} in order to reach Solar metallicity at the end of the model.

@@

We run {\tt VICE} in ``infall mode,'' where we specify the gas infall density $\dot\Sigma_{\rm in}$ and the star formation efficiency (SFE) timescale $\tau_\star\equiv \Sigma_g / \dot\Sigma_\star$ as functions of time. The gas surface density $\Sigma_g$ and star formation rate $\dot\Sigma_\star$ are calculated from the two specified quantities according to our star formation law, which is described in Section \ref{sec:sf-law}, assuming zero initial gas mass in all zones.

@@

Panel (b) of Figure \ref{fig:twoinfall-parameters} shows that 
% the second infall timescale $\tau_2$ is the most important parameter to tune in order to reproduce the observed stellar abundance distributions. 
the second infall timescale $\tau_2$ controls the size of the low-$\alpha$ loop, which affects the width of the MDF and the low-$\alpha$ [O/Fe] distribution. A shorter $\tau_2$ produces a bigger loop and therefore a broader [O/Fe] distribution which is skewed to higher [O/Fe], while a longer $\tau_2$ produces a smaller loop, leading to both a narrower low-$\alpha$ sequence and a narrower MDF. We note that our maximum value of $\tau_2=30\,{\rm Gyr}$ is close to a constant infall rate, so a further increase in $\tau_2$ has diminishing returns. Between $\tau_2=3-30\,{\rm Gyr}$, the endpoint of the abundance tracks shifts by $\sim0.2$ dex in [Fe/H] and $\sim0.1$ dex in [O/Fe], which could affect the model's ability to reproduce the present-day abundance of the Solar neighborhood. We adopt a fiducial value of $\tau_2=15\,{\rm Gyr}$ for the Solar neighborhood in order to minimize the size of the loop and width of the low-$\alpha$ [O/Fe] distribution while still approaching Solar [Fe/H] at late times (see further discussion in Section \ref{sec:abundance-distributions}). This value is in line with the infall timescale recovered by \citet{spitoni_galactic_2020}, and similar to the local star formation timescale adopted by \citet{johnson_stellar_2021}, but significantly longer than the timescales found by \citet{nissen_high-precision_2020} and \citet{spitoni_apogee_2021}. 
% In \todo{Section X}, we explore the effect of varying $\tau_2$ with radius in multi-zone models.

@@

We note that all our models are normalized to produce the same thick-to-thin-disk mass ratio of $f_{\Sigma}(R_\odot)=0.12$ \citep{bland-hawthorn_galaxy_2016} at the Solar annulus regardless of the infall parameters. The high-$\alpha$ sequence appears much less prominent in our [O/Fe] distributions in Figure \ref{fig:twoinfall-parameters} than in the data because the model outputs include only stars which were formed in-situ at the Solar annulus. In our multi-zone models, most of the high-$\alpha$ stars present in the Solar neighborhood have migrated from the inner Galaxy.

@@

\begin{figure*}
    \centering
    \includegraphics{figures/abundance_evolution_params.pdf}
    \caption{Stellar age--abundance relations in the Solar annulus produced by select multi-zone models with \yZ{1}. The layout is similar to Figure \ref{fig:abundance-evolution-yields}. Each column shows results from a different multi-zone model: {\bf (a)} our fiducial model with $y/Z_\odot=1$, $\sigma_{\rm RM8}=2.7\kpc$, pristine gas infall, and $f_\Sigma(R_\odot)=0.12$; {\bf (b)} a model with greater radial migration strength $\sigma_{\rm RM8}=5\kpc$; {\bf (c)} a model that assumes the infalling gas has metallicity $\mathOH_{\rm CGM}=\mathFeH_{\rm CGM}=-0.5$; and {\bf (d)} a model with a higher local thick-to-thin disk ratio, $f_\Sigma(R_\odot)=0.5$. {\bf Key takeaway:} Model (d) comes the closest to the observed age--metallicity relation at the expense of the age--[O/Fe] relation, but no model completely reconciles the dilution problem.}
    \label{fig:abundance-evolution-params}
\end{figure*}
